A Horrocks-type theorem for even orthogonal $K_2$

### 2020-07-10 horrocks.tex
Fix couple of defects

\begin{document}
%%%%% ------------- fill in your data below this line  -------------------
%%%%%    The following lines \Title ... \EndAddress must ALL be present
%%%%%    and in the given order.
\Title
A Horrocks-Type Theorem for Even Orthogonal $\mathrm{K}_2$
\ShortTitle 
A Horrocks-Type Theorem for $\mathrm{KO}_2$ 
\SubTitle   
%%%%%    A possible subtitle goes here.
\Author 
Andrei Lavrenov and Sergey Sinchuk
\ShortAuthor 
A. Lavrenov, S. Sinchuk
\EndTitle
\Abstract 
We prove the Horrocks theorem for unstable even-dimensional orthogonal Steinberg groups. The Horrocks theorem for Steinberg groups is one of the principal ingredients needed for the proof of the $\K_2$-analogue of Serre's problem, positive solution of which currently is only known in the linear case.
\EndAbstract
\MSC 
19C20
\EndMSC
\KEY 
Steinberg group, $\mathrm{K}_2$-functor, Quillen--Suslin theorem, Horrocks theorem, $\mathbb{P}^1$--glueing
\EndKEY
%%%%%    All 4 \Address lines below must be present. To center the last
%%%%%    entry, no empty lines must be between the following \Address
%%%%%    and \EndAddress lines.
\Address
Andrei Lavrenov
Chebyshev Laboratory 
St. Petersburg State University 
14th. Line V.O. 29b
199178, St. Petersburg
Russia
avlavrenov@gmail.com
\Address
Sergey Sinchuk
Chebyshev Laboratory 
St. Petersburg State University 
14th. Line V.O. 29b
199178, St. Petersburg
Russia
sinchukss@gmail.com
\Address
\Address
\EndAddress
%%
%%       Make sure the last tex command in your manuscript
%%       before the first \end{document} is the command  \Addresses
%%
%%---------------------Here the prologue ends---------------------------------
%%--------------------Here the manuscript starts------------------------------

@@

\section{Preliminaries}
If $x$ and $y$ are elements of a group $G$ then $[x, y]$ denotes the left-normed commutator $xyx^{-1}y^{-1}$.
We denote by $x^y$ (resp. ${}^y\!x$) the element $y^{-1}xy$ (resp. $yxy^{-1}$). 
Throughout the paper we make use of the following commutator identities:
\begin{align}
 [x, yz]^y =& [y^{-1}, x] \cdot [x, z], \label{rel43} \\ 
 \label{eq:H1ii} [xy, z] =& {}^x[y, z] \cdot [x,z], \\
 \label{eq:H1ii-2}[x, yz] =& [x, y] \cdot {}^{y}\![x, z], \\
 [[x, y], z] =& {}^{x}\!\left({}^{y}[x^{-1}, [ y^{-1}, z]] \cdot {}^z[y, [ z^{-1}, x^{-1}]] \right). \label{HW-variant} \end{align}
Recall also from~\cite[Lemma~3.1.1]{RS76} that
\begin{equation} \label{eq:H1iii} [x,z] = 1 \text{ implies } [x, [y,z]] = [[x,y],{}^yz]. \end{equation}

@@

In this paper we will be mostly interested in the case when the Dynkin diagram of $\Phi$ is {\it simply-laced}, i.\,e. does not contain double bonds. In this case the defining relations of $\St(\Phi, R)$ have the following simpler form:
\begin{align}
&x_{\alpha}(a)\cdot x_{\alpha}(b)=x_{\alpha}(a+b), \tag{R1} \label{Steinberg-additivity}\\
&[x_{\alpha}(a),\,x_{\beta}(b)]=x_{\alpha+\beta}(N_{\alpha\beta} \cdot ab),\text{ for }\alpha+\beta\in\Phi, \tag{R2} \label{Chevalley-CCF1} \\
&[x_{\alpha}(a),\,x_{\beta}(b)]=1,\text{ for }\alpha+\beta\not\in\Phi\cup0. \tag{R3} \label{Chevalley-CCF2}
\end{align}
In the above formulae $a, b \in R$ and the integers $N_{\alpha, \beta} = N_{\alpha, \beta, 1, 1} = \pm 1$ are the structure constants of the Lie algebra of type $\Phi$. Although there is still some degree of freedom in their choice, they always must satisfy the relations, indicated in the following lemma (cf. \cite[\S~14]{VP}).
\begin{lemma} Suppose $\Phi$ is simply laced and $\alpha, \beta$ are roots of $\Phi$ such that $\alpha+\beta\in \Phi$, then the following identities hold for the structure constants:
\begin{equation} \label{eq:simplest} N_{\alpha, \beta} = -N_{\beta,\alpha} = - N_{-\alpha, -\beta} = N_{\beta, -\alpha-\beta} = N_{-\alpha-\beta, \alpha}. \end{equation}
If, moreover, $\gamma \in \Phi$ is such that $\alpha,\beta,\gamma$ form a basis of a root subsystem of type $\rA_3$ then one has
\begin{equation} \label{eq:cocycle} N_{\beta,\gamma} \cdot N_{\alpha, \beta+\gamma} = N_{\alpha+\beta, \gamma} \cdot N_{\alpha, \beta}. \end{equation} \end{lemma}
In our calculations below we repeatedly use identities~\eqref{eq:simplest} without explicit reference.

@@

Recall that the classical Matsumoto theorem (see~\cite[Theorem~5.10]{Ma69}) allows one to compute the group $\K_2(\Phi, R)$ in the special case when $R=k$ is a field.
Using the modern language of Milnor--Witt K-theory (see~\cite{Mo04}) it can be formulated as follows:
\begin{equation*} \K_2(\Phi, k) = \left\{\begin{array}{ll} \K_2^\mathrm{MW}(k)& \text{if $\Phi$ is symplectic,}\\ \K_2^\mathrm{M}(k) & \text{otherwise.}\end{array}\right. \end{equation*}
In the following lemma we recall the computation of the group $\K_2(\Phi, R)$ in the case $R=k[X, X\inv]$.
\begin{lemma}[Hurrelbrink--Morita--Rehmann]\label{K2-laurent-field} Let $\Phi$ be a reduced irreducible root system of type $\neq \rG_2$ and $k$ be an arbitrary field. Then there is a split exact sequence of abelian groups
\[ \xymatrix{ 0 \ar[r] & \K_2(\Phi, k) \ar[r] & \K_2(\Phi, k[X, X^{-1}]) \ar[r] & H(\Phi, k) \ar[r] & 0,}\text{ in which} \]
\[ H(\Phi, k) = \left\{\begin{array}{ll} \K_1^\mathrm{MW}(k)& \text{if $\Phi$ is symplectic,}\\ \K_1(k) \cong k^\times & \text{otherwise.}  \end{array}\right. \]  \end{lemma}
\begin{proof} Let us first consider the case of nonsymplectic $\Phi$, in which one can find a long root $\alpha\in \Phi$ in such a way that there is a commutative diagram of abelian groups
\[\xymatrixcolsep{5pc}\xymatrix{k^\times \ar[r]^-{h} \ar[rrd]_{\{-, X\}} & \K_2(\Phi, k[X, X^{-1}]) \ar[r] & \K_2(\Phi, k(X)) \ar[d]^{\cong} \\
                                                                                      &                                 & \K_2^\mathrm{M}(k(X)),} \]
in which $h = \{ -, X \}_{\alpha}$ and the vertical map is an isomorphism by Matsumoto's theorem.
Notice that the diagonal map is split by the obvious residue homomorphism and therefore is injective. This, in turn, implies that $h$ is also injective.
The assertion of the lemma now follows from~\cite[Satz~3]{Hur77} which asserts that for a nonsymplectic $\Phi$ one has $\K_2(\Phi, k[X, X\inv]) = \mathrm{Im}(h) \oplus \K_2(\Phi, k)$.

@@

Recall that {\it the double ring} $D_{R, I}$ of a pair $(R, I)$ is, by definition, the pullback ring $R \times_{R/I} R$. In other words, it is the ring consisting of pairs of elements of $R$ congruent modulo $I$. Denote by $p_0, p_1, \Delta$ the two obvious projections and the diagonal map $\xymatrix{D_{R,I} \ar@<1pt>[r] \ar@<-2.5pt>[r] & \ar@<-4pt>[l] R}$. It is clear that $p_0 \Delta = p_1 \Delta = id_{R}$.

@@

\subsection{Computation of the kernel of the map of evaluation at 0.}\label{sec:computationOfK}
Let $A$ be a local ring with maximal ideal $M$.
The aim of this subsection is to describe a generating set for the kernel of the map $ev_{X=0}^*\colon\overline{\St}(\Phi, A[X], M[X]) \to \overline{\St}(\Phi, A, M)$
induced by the ring homomorphism of evaluation at $0$. We denote this kernel by $K(A[X], M[X])$.

@@

\subsection{The subgroups \texorpdfstring{$P_\alpha(0)$}{Pa(0)}, \texorpdfstring{$P_\alpha(*)$}{Pa(*)}, \texorpdfstring{$K(\alpha, \beta)$}{K(a,b)} and their properties.} \label{sec:Pa0-basic}
Let $\Phi$ be a root system and $\alpha$ be an element of $\Phi$.
Consider the following subsets of $\Phi$:
\begin{align} Z_+(\alpha) & = \{ \beta \in \Phi \mid \langle \alpha, \beta \rangle > 0 \}, \\
   Z_0(\alpha) & = \{ \beta \in \Phi \mid \alpha + \beta \not\in \Phi,\ \langle \alpha, \beta \rangle = 0 \}, \\
   Z(\alpha)   & = Z_0(\alpha) \sqcup Z_+(\alpha). \end{align}
Clearly, $Z_0(\alpha)$ (resp. $Z_+(\alpha)$) is a symmetric (resp. special) subset of $\Phi$ (see~\cref{df:root-subsets} for the terminology).

@@

\begin{rem} \label{rem:palpha} It follows from~\cref{P0_normal} and~\cref{Pstar-large} that the value of the function $p_\alpha$ from the statement of~\cref{P0_normal} on an element $g\in K(A[X], M[X])$ can be computed via the following procedure. 
Start with any presentation of $g$ as a product of elements $z_{\beta}(Xf, \xi)$ for $\beta \in \Phi$ and $c_\delta(f, X\xi)$ for some fixed $\delta\in Z_0(\alpha)$. 
Now pick among these factors those that correspond to the root $\beta = -\alpha$ (i.\,e. pick all factors $z_{-\alpha}(Xf_i, \xi_i)$). 
Now $p_\alpha(g)$ is precisely the sum of constant terms of the polynomials $f_i$. \end{rem}

@@

Now pick each factor $x_{\gamma}(Xf)$, $\gamma \in \{ -\alpha, -\beta \}$ appearing in the presentation for $g_2$, decompose it as $x_{\gamma}(X^2f') \cdot x_{\gamma}(m'X)$ for some $f' \in M[X]$, $m' \in M$ and then move the second factor to the rightmost position within $g_2$ conjugating all factors along the way. By the previous paragraph only elements of  $\mathcal{G}_5$ may appear modulo $G_0$ after simplification of these conjugates.
Thus, we can rewrite $g_2$ as $g_{21} \cdot g_{22}$, where $g_{21} \in \langle \mathcal{G}_2, \mathcal{G}_3, \mathcal{G}_4, \mathcal{G}_5, G_0 \rangle \subseteq \widetilde{K}(\alpha, \beta)$ and $g_{22} \in \langle X_{-\alpha}(MX), X_{-\beta}(MX) \rangle = X_{-\alpha}(MX) \cdot X_{-\beta}(MX)$.
Thus, we have obtained the desired decomposition for $g$. \end{proof}

@@

\subsection{The map \texorpdfstring{$S_\alpha(a, -)$}{S(a,-)} and its properties.} \label{sec:S-a}
Throughout this section $(A, M)$ is a local pair and, as before, $\alpha, \beta$ are some fixed roots of $\Phi$ forming a sharp angle.

@@

\subsection{Construction of a \texorpdfstring{$\St(\Phi, B)$}{St(B)}-torsor.} \label{sec:V-construction}
Throughout this subsection $\Phi$ denotes an arbitrary irreducible simply-laced root system of rank $\geq 3$, unless stated otherwise. 
\begin{df}\label{df:GM0_geq0}
Denote the subgroup $\overline{\St}(\Phi, A, M)$ by $G_M^0$. 
Consider the following commutative diagram, in which the homomorphisms $i_\pm$ and $j_\pm$ are induced by the natural ring homomorphisms $A \to A[X^\pm]$ and $A[X^\pm]\to R$:
\newline \scalebox{0.80}{\begin{minipage}{1.15\linewidth}
\[\xymatrix{ G_M^0 \ar@{^{(}->}[rr]^{i_+^M} \ar@<-5pt>@{^{(}->}[dd]^{i_-^M} \ar@<-5pt>@{^{(}->}[rd] & & \overline{\St}(\Phi, A[X], M[X]) \ar[dd]_(.3){j_+^M} \ar@<-2.5pt>@{^{(}->}[rd] & \\ & \St(\Phi, A) \ar@{^{(}->}[rr]^(.3){i_+} \ar@<-2.5pt>@{^{(}->}[dd]^(.3){i_-} & & \St(\Phi, A[X]) \ar[dd]^{j_+} \\ \overline{\St}(\Phi, A[X\inv], M[X\inv]) \ar@<-2.5pt>@{^{(}->}[rd] \ar[rr]^(.35){j_-^M} & & \overline{\St}(\Phi, R, I) \ar@<-2.5pt>@{^{(}->}[rd] & \\ & \St(\Phi, A[X\inv]) \ar[rr]_{j_-} & & \St(\Phi, R).} \]\end{minipage}} \vskip \belowdisplayshortskip
Denote by $\overline{G}^{\geq 0}_M$ the image of the homomorphism $j_+^M$. \end{df}
Notice that the homomorphism $j_+^Mi_+^M$ is split by the homomorphism $ev^*_{X=1}$ of evaluation at $1$, therefore $G_M^0$ can be considered as a subgroup of $\overline{G}^{\geq 0}_M$.

@@

\subsection{Proof of Horrocks theorem.} \label{sec:P1glueing}
Before we proceed with the proof of~\cref{thm:P1glueing} let us briefly recall the relevant notation. As before, $A$ denotes an arbitary local ring with maximal ideal $M$ and $I = M[X, X\inv]$ is an ideal of both the ring $R=A[X, X\inv]$ and its subring $B = A[X\inv] + M[X]$.

@@

\begin{thebibliography}{1000}
\bibitem{Abe83}
 E.~Abe, 
 \emph{\href{https://dx.doi.org/10.1080/00927878308822906} 
 {Whitehead groups of {C}hevalley groups over polynomial rings}},
 Comm. Alg. \textbf{11} (1983), no.~12, 1271--1307.

\bibitem{A13}
 D.~Allcock, 
 \emph{\href{https://dx.doi.org/10.2140/ant.2016.10.1791} 
 {Steinberg groups as amalgams}},
 Alg. Number Th. \textbf{10} (2016), no.~8, 1791--1843.

\bibitem{AF17}
 A.~Asok and J.~Fasel, 
 \emph{\href{https://dx.doi.org/10.1112/topo.12007} 
 {An explicit KO-degree map and applications}},
 J. Top. \textbf{10} (2017), 268--300.

\bibitem{Ba74}
 H.~Bass, 
 \emph{\href{https://dx.doi.org/10.2307/2373586} 
 {Clifford algebras and spinor norms over a commutative ring}},
 Amer. J. Math. \textbf{96} (1974), no.~1, 156--206.

@@

\bibitem{DS73}
 R.~K.~Dennis and M.~R.~Stein,
 \emph{\href{https://dx.doi.org/10.1007/BFb0073728} 
 {The functor $\K_2$: A survey of computations and problems}},
 Classical Algebraic K-Theory, and Connections with Arithmetic, Springer, 1973, pp.~241--303.

\bibitem{FRS12}
 J.~Fasel, R.~Rao, and R.~Swan,
 \emph{\href{https://dx.doi.org/10.1007/s10240-012-0041-y} 
 {On stably free modules over affine algebras}},
 Publ. Math. IH{\'E}S \textbf{116} (2012), 223--243.

\bibitem{Ho05}
 J.~Hornbostel, 
 \emph{\href{https://dx.doi.org/10.1016/j.top.2004.10.004} 
 {$\mathbb{A}^1$-representability of hermitian K-theory and Witt groups}},
 J. Top. \textbf{44} (2005), no.~3, 661--687.

\bibitem{Ho64}
 G.~Horrocks, 
 \emph{\href{https://dx.doi.org/10.1112/plms/s3-14.4.714} 
 {Projective modules over an extension of a local ring}},
 Proc. London Math. Soc. \textbf{3} (1964), no.~4, 714--718.

\bibitem{Hur77}
 J.~Hurrelbrink, 
 \emph{\href{https://dx.doi.org/10.1007/BF01351716} 
 {Endlich pr\"asentierte arithmetische gruppen und $K_{2}$ {\"u}ber Laurent-polynomringen}},
 Math. Ann. \textbf{225} (1977), no.~2, 123--129.

\bibitem{Ka80}
 M.~Karoubi, 
 \emph{\href{https://dx.doi.org/10.2307/1971147} 
 {Le th{\'e}oreme fondamental de la K-th{\'e}orie hermitienne}},
 Ann. Math. \textbf{112} (1980), no.~2, 259--282.

@@

\bibitem{La18}
 A.~Lavrenov, 
 \emph{\href{https://dx.doi.org/10.25537/dm.2018v23.653-675}
 {A local-global principle for symplectic $\K_2$}},
 Doc. Math. \textbf{23} (2018), 653--675.

\bibitem{LS17}
 A.~Lavrenov and S.~Sinchuk,
 \emph{\href{https://arxiv.org/pdf/1606.00272.pdf} 
 {Centrality of odd orthogonal $\K_2$}},
 J. Pure Appl. Alg. \textbf{221} (2017), 1134--1145.

\bibitem{Lo78}
 J.-L.~Loday,
 \emph{\href{https://dx.doi.org/10.1016/0021-8693(78)90025-X}
 {Cohomologie et groupes de Steinberg relatifs}}, 
 J. Alg. \textbf{54} (1978), no.~1, 178--202.

\bibitem{Ma69}
 H.~Matsumoto,
 \emph{\href{http://eudml.org/doc/81843} 
 {Sur les sous-groupes arithm{\'e}tiques des groupes semi-simple d{\'e}ploy{\'e}s}},
 Ann. Sci. ENS \textbf{2} (1969), no.~1, 1--62.

>>>>

\begin{document}
%%%%% ------------- fill in your data below this line  -------------------
%%%%%    The following lines \Title ... \EndAddress must ALL be present
%%%%%    and in the given order.
\Title
A Horrocks-Type Theorem for Even Orthogonal $\mathrm{K}_2$
\ShortTitle 
A Horrocks-Type Theorem for $\mathrm{KO}_2$ 
\SubTitle   
%%%%%    A possible subtitle goes here.
\Author 
Andrei Lavrenov and Sergey Sinchuk
\ShortAuthor 
A. Lavrenov, S. Sinchuk
\EndTitle
\Abstract 
We prove the Horrocks theorem for unstable even-dimensional orthogonal Steinberg groups. The Horrocks theorem for Steinberg groups is one of the principal ingredients needed for the proof of the $\K_2$-analogue of Serre’s problem,  whose positive solution is currently known only in the linear case.
\EndAbstract
\MSC 
19C20
\EndMSC
\KEY 
Steinberg group, $\mathrm{K}_2$-functor, Quillen--Suslin theorem, Horrocks theorem, $\mathbb{P}^1$--glueing
\EndKEY
%%%%%    All 4 \Address lines below must be present. To center the last
%%%%%    entry, no empty lines must be between the following \Address
%%%%%    and \EndAddress lines.
\Address
Andrei Lavrenov
Chebyshev Laboratory 
St. Petersburg State University 
14th. Line V.O. 29b
199178, St. Petersburg
Russia
avlavrenov@gmail.com
\Address
Sergey Sinchuk
Chebyshev Laboratory 
St. Petersburg State University 
14th. Line V.O. 29b
199178, St. Petersburg
Russia
sinchukss@gmail.com
\Address
\Address
\EndAddress
%%
%%       Make sure the last tex command in your manuscript
%%       before the first \end{document} is the command  \Addresses
%%
%%---------------------Here the prologue ends---------------------------------
%%--------------------Here the manuscript starts------------------------------

@@

\section{Preliminaries}
If $x$ and $y$ are elements of a group $G$ then $[x, y]$ denotes the left-normed commutator $xyx^{-1}y^{-1}$.
We denote by $x^y$ (resp. ${}^y\!x$) the element $y^{-1}xy$ (resp. $yxy^{-1}$). 
Throughout the paper we make use of the following commutator identities:
\begin{align}
 [x, yz]^y               & =   [y^{-1}, x] \cdot [x, z], \label{rel43} \\ 
 \label{eq:H1ii} [xy, z] & =   {}^x[y, z] \cdot [x,z], \\
 \label{eq:H1ii-2}[x, yz]& =   [x, y] \cdot {}^{y}\![x, z], \\
 [[x, y], z] & =   {}^{x}\!\left({}^{y}[x^{-1}, [ y^{-1}, z]] \cdot {}^z[y, [ z^{-1}, x^{-1}]] \right). \label{HW-variant} \end{align}
Recall also from~\cite[Lemma~3.1.1]{RS76} that
\begin{equation} \label{eq:H1iii} [x,z] = 1 \text{ implies } [x, [y,z]] = [[x,y],{}^yz]. \end{equation}

@@

In this paper we will be mostly interested in the case when the Dynkin diagram of $\Phi$ is {\it simply-laced}, i.\,e. does not contain double bonds. In this case the defining relations of $\St(\Phi, R)$ have the following simpler form:
\begin{align}
x_{\alpha}(a)\cdot x_{\alpha}(b)&=x_{\alpha}(a+b), \tag{R1} \label{Steinberg-additivity}\\
[x_{\alpha}(a),\,x_{\beta}(b)]  &=x_{\alpha+\beta}(N_{\alpha\beta} \cdot ab),\text{ for }\alpha+\beta\in\Phi, \tag{R2} \label{Chevalley-CCF1} \\
[x_{\alpha}(a),\,x_{\beta}(b)]  &=1,\text{ for }\alpha+\beta\not\in\Phi\cup0. \tag{R3} \label{Chevalley-CCF2}
\end{align}
In the above formulae $a, b \in R$ and the integers $N_{\alpha, \beta} = N_{\alpha, \beta, 1, 1} = \pm 1$ are the structure constants of the Lie algebra of type $\Phi$. Although there is still some degree of freedom in their choice, they always must satisfy the relations, indicated in the following lemma (cf. \cite[\S~14]{VP}).
\begin{lemma} Suppose $\Phi$ is simply laced and $\alpha, \beta$ are roots of $\Phi$ such that $\alpha+\beta\in \Phi$, then the following identities hold for the structure constants:
\begin{equation} \label{eq:simplest} N_{\alpha, \beta} = -N_{\beta,\alpha} = - N_{-\alpha, -\beta} = N_{\beta, -\alpha-\beta} = N_{-\alpha-\beta, \alpha}. \end{equation}
If, moreover, $\gamma \in \Phi$ is such that $\alpha,\beta,\gamma$ form a basis of a root subsystem of type $\rA_3$ then one has
\begin{equation} \label{eq:cocycle} N_{\beta,\gamma} \cdot N_{\alpha, \beta+\gamma} = N_{\alpha+\beta, \gamma} \cdot N_{\alpha, \beta}. \end{equation} \end{lemma}
In our calculations below we repeatedly use identities~\eqref{eq:simplest} without explicit reference.

@@

Recall that the classical Matsumoto theorem (see~\cite[Theorem~5.10]{Ma69}) allows one to compute the group $\K_2(\Phi, R)$ in the special case when $R=k$ is a field.
Using the modern language of Milnor--Witt K-theory (see~\cite{Mo04}) it can be formulated as follows:
\begin{equation*} \K_2(\Phi, k) = \left\{\begin{array}{ll} \K_2^\mathrm{MW}(k)& \text{if $\Phi$ is symplectic,}\\ \K_2^\mathrm{M}(k) & \text{otherwise.}\end{array}\right. \end{equation*}
In the following lemma we recall the computation of the group $\K_2(\Phi, R)$ in the case $R=k[X, X\inv]$.
\begin{lemma}[Hurrelbrink--Morita--Rehmann]\label{K2-laurent-field} Let $\Phi$ be a reduced irreducible root system of type $\neq \rG_2$ and $k$ be an arbitrary field. Then there is a split exact sequence of abelian groups
\[ \xymatrix{ 0 \ar[r] & \K_2(\Phi, k) \ar[r] & \K_2(\Phi, k[X, X^{-1}]) \ar[r] & H(\Phi, k) \ar[r] & 0,}\text{ in which} \]
\[ H(\Phi, k) = \left\{\begin{array}{ll} \K_1^\mathrm{MW}(k)& \text{if $\Phi$ is symplectic,}\\ \K_1(k) \cong k^\times & \text{otherwise.}  \end{array}\right. \]  \end{lemma}
\begin{proof} Let us first consider the case of nonsymplectic $\Phi$, in which one can find a long root $\alpha\in \Phi$ in such a way that there is a commutative diagram of abelian groups
\[\xymatrixcolsep{5pc}\xymatrix{k^\times \ar[drr]_{\{-, X\}} \ar[r]^-{h}  & \K_2(\Phi, k[X, X^{-1}]) \ar[r] & \K_2(\Phi, k(X)) \ar[d]^{\cong} \\
                                                                                      &                                 & \K_2^\mathrm{M}(k(X)),} \]
in which $h = \{ -, X \}_{\alpha}$ and the vertical map is an isomorphism by Matsumoto's theorem.
Notice that the diagonal map is split by the obvious residue homomorphism and therefore is injective. This, in turn, implies that $h$ is also injective.
The assertion of the lemma now follows from~\cite[Satz~3]{Hur77} which asserts that for a nonsymplectic $\Phi$ one has $\K_2(\Phi, k[X, X\inv]) = \mathrm{Im}(h) \oplus \K_2(\Phi, k)$.

@@

Recall that {\it the double ring} $D_{R, I}$ of a pair $(R, I)$ is, by definition, the pullback ring $R \times_{R/I} R$. In other words, it is the ring consisting of pairs of elements of~$R$ congruent modulo $I$. Denote by $p_0, p_1, \Delta$ the two obvious projections and the diagonal map $\xymatrix{D_{R,I} \ar@<1pt>[r] \ar@<-2.5pt>[r] & \ar@<-4pt>[l] R}$. It is clear that $p_0 \Delta = p_1 \Delta = id_{R}$.

@@

\subsection{Computation of the kernel of the map of evaluation at 0}\label{sec:computationOfK}
Let $A$ be a local ring with maximal ideal $M$.
The aim of this subsection is to describe a generating set for the kernel of the map $ev_{X=0}^*\colon\overline{\St}(\Phi, A[X], M[X]) \to \overline{\St}(\Phi, A, M)$
induced by the ring homomorphism of evaluation at $0$. We denote this kernel by $K(A[X], M[X])$.

@@

\subsection{The subgroups \texorpdfstring{$P_\alpha(0)$}{Pa(0)}, \texorpdfstring{$P_\alpha(*)$}{Pa(*)}, \texorpdfstring{$K(\alpha, \beta)$}{K(a,b)} and their properties} \label{sec:Pa0-basic}
Let $\Phi$ be a root system and $\alpha$ be an element of $\Phi$.
Consider the following subsets of $\Phi$:
\begin{align} Z_+(\alpha) & = \{ \beta \in \Phi \mid \langle \alpha, \beta \rangle > 0 \}, \\
   Z_0(\alpha) & = \{ \beta \in \Phi \mid \alpha + \beta \not\in \Phi,\ \langle \alpha, \beta \rangle = 0 \}, \\
   Z(\alpha)   & = Z_0(\alpha) \sqcup Z_+(\alpha). \end{align}
Clearly, $Z_0(\alpha)$ (resp. $Z_+(\alpha)$) is a symmetric (resp. special) subset of $\Phi$ (see~\cref{df:root-subsets} for the terminology).

@@

\begin{rem} \label{rem:palpha} It follows from~\cref{P0_normal} and~\cref{Pstar-large} that the value of the function $p_\alpha$ from the statement of~\cref{P0_normal} on an element $g\in K(A[X], M[X])$ can be computed via the following procedure. 
Start with any presentation of~$g$ as a product of elements $z_{\beta}(Xf, \xi)$ for $\beta \in \Phi$ and $c_\delta(f, X\xi)$ for some fixed $\delta\in Z_0(\alpha)$. 
Now pick among these factors those that correspond to the root $\beta = -\alpha$ (i.\,e. pick all factors $z_{-\alpha}(Xf_i, \xi_i)$). 
Now $p_\alpha(g)$ is precisely the sum of constant terms of the polynomials $f_i$. \end{rem}

@@

Now pick each factor $x_{\gamma}(Xf)$, $\gamma \in \{ -\alpha, -\beta \}$ appearing in the presentation for~$g_2$, decompose it as $x_{\gamma}(X^2f') \cdot x_{\gamma}(m'X)$ for some $f' \in M[X]$, $m' \in M$ and then move the second factor to the rightmost position within $g_2$ conjugating all factors along the way. By the previous paragraph only elements of  $\mathcal{G}_5$ may appear modulo $G_0$ after simplification of these conjugates.
Thus, we can rewrite $g_2$ as $g_{21} \cdot g_{22}$, where $g_{21} \in \langle \mathcal{G}_2, \mathcal{G}_3, \mathcal{G}_4, \mathcal{G}_5, G_0 \rangle \subseteq \widetilde{K}(\alpha, \beta)$ and $g_{22} \in \langle X_{-\alpha}(MX), X_{-\beta}(MX) \rangle = X_{-\alpha}(MX) \cdot X_{-\beta}(MX)$.
Thus, we have obtained the desired decomposition for $g$. \end{proof}

@@

\subsection{The map \texorpdfstring{$S_\alpha(a, -)$}{S(a,-)} and its properties} \label{sec:S-a}
Throughout this section $(A, M)$ is a local pair and, as before, $\alpha, \beta$ are some fixed roots of $\Phi$ forming a sharp angle.

@@

\subsection{Construction of a \texorpdfstring{$\St(\Phi, B)$}{St(B)}-torsor} \label{sec:V-construction}
Throughout this subsection $\Phi$ denotes an arbitrary irreducible simply-laced root system of rank $\geq 3$, unless stated otherwise. 
\begin{df}\label{df:GM0_geq0}
Denote the subgroup $\overline{\St}(\Phi, A, M)$ by $G_M^0$. 
Consider the following commutative diagram, in which the homomorphisms $i_\pm$ and $j_\pm$ are induced by the natural ring homomorphisms $A \to A[X^\pm]$ and $A[X^\pm]\to R$:
\newline \scalebox{0.80}{\begin{minipage}{1.15\linewidth}
\[\xymatrix{ G_M^0 \ar@{^{(}->}[rr]^{i_+^M} \ar@<-5pt>@{^{(}->}[dd]^{i_-^M} \ar@<-5pt>@{^{(}->}[rd] & & \overline{\St}(\Phi, A[X], M[X]) \ar[dd]_(.3){j_+^M} \ar@<-2.5pt>@{^{(}->}[rd] & \\ & \St(\Phi, A) \ar@{^{(}->}[rr]^(.3){i_+} \ar@<-2.5pt>@{^{(}->}[dd]^(.3){i_-} & & \St(\Phi, A[X]) \ar[dd]^{j_+} \\ \overline{\St}(\Phi, A[X\inv], M[X\inv]) \ar@<-2.5pt>@{^{(}->}[rd] \ar[rr]^(.35){j_-^M} & & \overline{\St}(\Phi, R, I) \ar@<-2.5pt>@{^{(}->}[rd] & \\ & \St(\Phi, A[X\inv]) \ar[rr]_{j_-} & & \St(\Phi, R).} \]\end{minipage}} \vskip \belowdisplayshortskip
Denote by $\overline{G}^{\geq 0}_M$ the image of the homomorphism $j_+^M$. \end{df}
Notice that the homomorphism $j_+^Mi_+^M$ is split by the homomorphism $ev^*_{X=1}$ of evaluation at $1$, therefore $G_M^0$ can be considered as a subgroup of $\overline{G}^{\geq 0}_M$.

@@

\subsection{Proof of Horrocks theorem} \label{sec:P1glueing}
Before we proceed with the proof of~\cref{thm:P1glueing} let us briefly recall the relevant notation. As before, $A$ denotes an arbitary local ring with maximal ideal $M$ and $I = M[X, X\inv]$ is an ideal of both the ring $R=A[X, X\inv]$ and its subring $B = A[X\inv] + M[X]$.

@@

\begin{thebibliography}{1000}
\bibitem{Abe83}
 E.~Abe, 
 \emph{Whitehead groups of {C}hevalley groups over polynomial rings},
 Comm. Alg. \textbf{11} (1983), no.~12, 1271--1307.

\bibitem{A13}
 D.~Allcock, 
 \emph{Steinberg groups as amalgams},
 Alg. Number Th. \textbf{10} (2016), no.~8, 1791--1843.

\bibitem{AF17}
 A.~Asok and J.~Fasel, 
 \emph{An explicit KO-degree map and applications},
 J. Top. \textbf{10} (2017), 268--300.

\bibitem{Ba74}
 H.~Bass, 
 \emph{Clifford algebras and spinor norms over a commutative ring},
 Amer. J. Math. \textbf{96} (1974), no.~1, 156--206.

@@

\bibitem{DS73}
 R.~K.~Dennis and M.~R.~Stein,
 \emph{The functor $\K_2$: A survey of computations and problems},
 Classical Algebraic K-Theory, and Connections with Arithmetic, Springer, 1973, pp.~241--303.

\bibitem{FRS12}
 J.~Fasel, R.~Rao, and R.~Swan,
 \emph{On stably free modules over affine algebras},
 Publ. Math. IH{\'E}S \textbf{116} (2012), 223--243.

\bibitem{Ho05}
 J.~Hornbostel, 
 \emph{$\mathbb{A}^1$-representability of hermitian K-theory and Witt groups},
 J. Top. \textbf{44} (2005), no.~3, 661--687.

\bibitem{Ho64}
 G.~Horrocks, 
 \emph{Projective modules over an extension of a local ring},
 Proc. London Math. Soc. \textbf{3} (1964), no.~4, 714--718.

\bibitem{Hur77}
 J.~Hurrelbrink, 
 \emph{Endlich pr\"asentierte arithmetische gruppen und $K_{2}$ {\"u}ber Laurent-polynomringen},
 Math. Ann. \textbf{225} (1977), no.~2, 123--129.

\bibitem{Ka80}
 M.~Karoubi, 
 \emph{Le th{\'e}oreme fondamental de la K-th{\'e}orie hermitienne},
 Ann. Math. \textbf{112} (1980), no.~2, 259--282.

@@

\bibitem{La18}
 A.~Lavrenov, 
 \emph{A local-global principle for symplectic $\K_2$},
 Doc. Math. \textbf{23} (2018), 653--675.

\bibitem{LS17}
 A.~Lavrenov and S.~Sinchuk,
 \emph{entrality of odd orthogonal $\K_2$},
 J. Pure Appl. Alg. \textbf{221} (2017), 1134--1145.

\bibitem{Lo78}
 J.-L.~Loday,
 \emph{Cohomologie et groupes de Steinberg relatifs}, 
 J. Alg. \textbf{54} (1978), no.~1, 178--202.

\bibitem{Ma69}
 H.~Matsumoto,
 \emph{Sur les sous-groupes arithm{\'e}tiques des groupes semi-simple d{\'e}ploy{\'e}s},
 Ann. Sci. ENS \textbf{2} (1969), no.~1, 1--62.

\bibitem{Mo04}
 F.~Morel,
 \emph{Sur les puissances de l'id{\'e}al fondamental de l'anneau de Witt},
 Comm. Math. Helv. \textbf{79} (2004), no.~4, 689--703.

\bibitem{MR91}
 J.~Morita and U.~Rehmann,
 \emph{Symplectic $\K_2$ of Laurent polynomials, associated Kac--Moody groups and Witt rings},
 Math. Z. \textbf{206} (1991), no.~1, 57--66.

\bibitem{Pa89}
 I.~Panin, 
 \emph{On stabilization for orthogonal and symplectic algebraic K-theory (Russian)},
 Leningrad Math. J. \textbf{1} (1990), no.~3, 741--754.
